---
title: "COVID19 in the Southeast"
author: "Brian Dill"
date: "01/20/2022"
output:
  html_document: 
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
    number_sections: no
    css: style_01.css    
params:
  state: Tennessee
  state_abbr: TN
  counties: !r c("Davidson County","Williamson County", "Rutherford County", "Shelby County", "Wilson County", "Moore County", "Meigs County" )

---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)  #https://haozhu233.github.io/kableExtra/awesome_table_in_html.html
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## Vaccination Rates by County

Data obtained from the CDC can be found on their [COVID-19 Vaccinations in the United States,County](https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh) page.  It is a large dataset of about 1M rows and 32 columns.  The CSV file is about 130MB.

This report is filtering to only data from `r params$state` and only selecting 6 columns.

```{r vax-load, message=FALSE}
#path <- "C:/Users/bdill/Downloads/COVID-19_Vaccinations_in_the_United_States_County_2021-10-05.csv"
path <- "C:/Users/bdill/Downloads/COVID-19_Vaccinations_in_the_United_States_County.csv"
#path <- "http://briandill.com/data/cdc/COVID-19_Vaccinations_in_the_United_States_County_2021-10-05.csv"

vax_raw <- data.table::fread(path)
vax <- vax_raw %>% 
  select(Date, FIPS, Recip_County, Recip_State, Series_Complete_Pop_Pct, Series_Complete_Yes) %>% 
  filter(Recip_State == params$state_abbr)

vax$Date <- lubridate::mdy(vax$Date)

head(vax) %>% kableExtra::kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```


## Vaccination Rate Chart

This shows the vaccination rate over time for a select number of counties in `r params$state`.  Williamson county was the most vaccinated until they were recently overtaken by Meigs County.  Moore county is included as it is the least vaccinated county.

```{r plot-vax}
vax %>% 
  filter(Recip_State == params$state_abbr, Recip_County %in% params$counties ) %>% 
  select(Date, FIPS, Recip_County, Recip_State, Series_Complete_Pop_Pct, Series_Complete_Yes) %>% 
  ggplot(aes(x = Date, y = Series_Complete_Pop_Pct, group = Recip_County)) +
  geom_line(aes(color = Recip_County), size = 1.2) +
  labs(title = paste("Total Vaccination Rate:", params$state,
       y = "Percent Fully Vaccinated"),
       subtitle = "Data: CDC",
       caption = "chart: @bdill\nData: CDC.  https://bit.ly/3oAMW7l")
```

## State Populations

State population totals from the US Census on their [State Population Totals: 2010-2020](https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-total.html) page.  

The direct [Excel file download link](https://www2.census.gov/programs-surveys/popest/tables/2010-2020/state/totals/nst-est2020.xlsx).  Or you can [download a CSV file](http://briandill.com/data/census/census_state_pop_nst-est2020.csv) directly from my website which is tiny (less than < 10kb)

The population estimate for 2020 for each state is being used.  Florida is the 4th largest US state and by far the largest in the southeast.

```{r state-pop }
pop <- read_csv("http://briandill.com/data/census/census_state_pop_nst-est2020.csv")
states <- read_csv("http://briandill.com/data/fips_state_usda.csv")

statepop <- states %>% 
  inner_join(pop, by = c("state_fips" = "FipsState")) %>% 
  select(state_fips, state, state_abbrev, pop = PopEst2020)

statepop %>% 
  filter(state_abbrev %in% c("TN", "MS", "GA", "FL", "AL", "LA")) %>% 
  arrange(desc(pop)) %>% 
  kableExtra::kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## COVID Cases by State Over Time

[CDC source page](https://data.cdc.gov/Case-Surveillance/United-States-COVID-19-Cases-and-Deaths-by-State-o/9mfq-cb36).  The CSV file is about 3MB ([direct download link](https://data.cdc.gov/api/views/9mfq-cb36/rows.csv?accessType=DOWNLOAD))

Note the extremely steep curves beginning in September 2021 (08/21).  This is the onslaught of the delta variant which is about twice as communicable as the original strain.

```{r case-death}
path <- "http://briandill.com/data/cdc/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv"

case_death <- data.table::fread(path)
case_death$submission_date <- lubridate::mdy(case_death$submission_date)
df <- case_death %>% 
  inner_join(statepop, by = c("state" = "state_abbrev")) %>% 
  mutate(tot_cases_per_100k = tot_cases / (pop / 100000),
         tot_deaths_per_100k = tot_death / (pop / 100000)) %>% 
  select(date = submission_date, state, pop, tot_cases, tot_death, tot_cases_per_100k, tot_deaths_per_100k)

df %>% filter(state %in% c("AL", "TN", "MS", "FL", "GA", "LA")) %>% 
  ggplot(aes(x = date, y = tot_cases_per_100k)) +
  geom_line(aes(color = state), size = 1) +
  scale_x_date(breaks = "2 months", date_labels = "%m/%y") +
  labs(title = "COVID 19 Cases by State Over Time",
       x = "Date",
       y = "Total Cases per 100K",
       caption = "chart: @bdill\ndata: CDC")

```

## COVID Deaths by State over Time

Using the same dataset and plotting deaths per 100k.  It is notable that while TN had the highest number of cases per 100k, it has the **lowest number of deaths** per 100k.

```{r plot-deaths}
df %>% filter(state %in% c("AL", "TN", "MS", "FL", "GA", "LA")) %>% 
  ggplot(aes(x = date, y = tot_deaths_per_100k)) +
  geom_line(aes(color = state), size = 0.9) +
  scale_x_date(breaks = "2 months", date_labels = "%m/%y") +
  labs(title = "COVID 19 Deaths by State Over Time",
       x = "Date",
       y = "Total Deaths per 100K",
       caption = "chart: @bdill\ndata: CDC")
```

