---
title: "Investment Report for Projects in `r params$country`"
author: "Brian Dill"
date: "Last edited: `r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
        collapsed: false
    number_sections: true
params:
    country: Turkey
    year_start: 2017-07-01
    year_end: 2018-06-30
    fy: 2018
---
<style>

#TOC {
  font-size: 16px;
}
#header {
  color: #800000;
}
body {
  color: #708090;
  font-family: Calibri;
  background-color: #F5F5F5;
}
pre {
  background-color: #F8F8FF;
}

</style>
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "80%", echo = TRUE, warning = FALSE)
```

```{r data, message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_region_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/52f5414f6504e0503e86eb1043afa9b3d157fab2/investment_region_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```


## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.
```{r investment-annual-summary, out.width="85%", fig.cap = "Figure 1.1 The Investment Annual Summary for each region for 2012 to 2018"}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

* East Asia and the Pacific
* Europe and Central Asia
* Latin America and the Caribbean
* Middle East and North Africa
* South Asia
* Sub-Saharan Africa

```{r tables, echo=FALSE}
kable(investment_region_summary, col.names = c("Region", "Dollars in Millions"), align = "cc", caption = "Table 1.1 The total investment summary for each region for the 2012 to 2018 fiscal years.")
```

### Investment Projects in `r params$country`

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.
```{r country-investment-projects, out.width="95%", fig.cap = paste0("Figure 1.2 The Investment Services Projects in ", params$country, " from 2012 to 2018"), collapse = TRUE}

country_investment_projects <- investment_services_projects %>%
  filter(country == params$country)

ggplot(country_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
    geom_point() +
    labs(
    title = "Investment Services Projects",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```

### Investment Projects in `r params$country` in `r params$fy`

```{r country-annual-investment-projects, out.width="95%", fig.cap = paste("Figure 1.3 The Investment Services Projects in", params$country, "in", params$fy), collapse = TRUE}

country_annual_investment_projects <- investment_services_projects %>%
  filter(country == params$country,
         date_disclosed >= params$year_start,
         date_disclosed <= params$year_end) 

#country_annual_investment_projects

ggplot(country_annual_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = paste0("Investment Services Projects in ", params$country, " in ", params$fy),
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  ) 
```

```{r}
country_annual_investment_projects_total <- country_annual_investment_projects %>%
  summarize(sum_total_investment = sum(total_investment, na.rm = TRUE)) 
```

The total investment amount for all projects in `r params$country` in the `r params$fy` fiscal year was `r country_annual_investment_projects_total` million dollars.