---
title: "US Census 2020"
author: "Brian Dill"
date: 'Last edited: `r format(Sys.time(), "%Y-%m-%d %H:%M")`'
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
    number_sections: no
    css: style_01.css

params:
  top_n: 10
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "80%", echo = FALSE, warning = FALSE)
```

# US Census 2020 

This report shows the top `r params$top_n` counties by highest percentage of a specific race.  The races in this report are the "alone" versions.  Ex: "White alone", "Black alone", etc.  For more detail, see the [Census Race Definitions](https://www.census.gov/programs-surveys/cps/data/data-tools/cps-table-creator-help/race-definitions.html)

Source data: US Census 2020 Decennial Data. Table P1 (Race) [https://data.census.gov/cedsci/table?q=race%202020&tid=DECENNIALPL2020.P1](https://data.census.gov/cedsci/table?q=race%202020&tid=DECENNIALPL2020.P1)

R markdown file used to create this report on [github](https://github.com/wbdill/r-sandbox01/blob/master/govt_data/census_2020_race.Rmd)


```{r data, echo = TRUE, message = FALSE}
library(tidyverse)
library(knitr)

path <- "http://briandill.com/data/census/DECENNIALPL2020.P1/DECENNIALPL2020.P1_data_with_overlays_2021-09-17T080619.csv"

df <- read_csv(path, skip = 2, col_names = FALSE)
df <- df[,1:11]
colnames(df) <- c("geoid", "name", "total", "one", "white", "black", "indian", 
                  "asian", "pacific", "other", "twoormore")
df <- separate(df, name, into = c("county", "state"), ", ")
df <- df %>% mutate(fips_county = substring(geoid, 10),
                    fips_state = substring(geoid, 10, 11),
                    white_pct = round(white*100 / total, 3),
                    black_pct = round(black*100 / total, 3),
                    indian_pct = round(indian*100 / total, 3),
                    asian_pct = round(asian*100 / total, 3),
                    pacific_pct = round(pacific*100 / total, 3),
                    other_pct = round(other*100 / total, 3),
                    twoormore_pct = round(twoormore*100 / total, 3) ) %>% 
  select(geoid, fips_state, fips_county, county, state, total, white, black, indian, asian, pacific
         , other, twoormore, one, white_pct, black_pct, indian_pct, asian_pct, pacific_pct
         , other_pct, twoormore_pct)
```


## Top `r params$top_n` Counties % White

Kentucky contains several of the top counties with the highest % of white population.

```{r echo = TRUE, message = FALSE}
df %>% arrange(desc(white_pct)) %>% 
  select(state, county, total, white, white_pct) %>% 
  head(params$top_n)%>% 
  kable()
```

## Top `r params$top_n` Counties % Black

Mississippi leads with the top counties with the highest % of black population.  This is the [black belt](https://en.wikipedia.org/wiki/Black_Belt_in_the_American_South)

```{r message = FALSE}
df %>% arrange(desc(black_pct)) %>% 
  select(state, county, total, black, black_pct) %>% 
  head(params$top_n) %>% 
  kable()
```

## Top `r params$top_n` Counties % American Indian or Alaska Native

```{r message = FALSE}
df %>% arrange(desc(indian_pct)) %>% 
  select(state, county, total, indian, indian_pct) %>% 
  head(params$top_n) %>% 
  kable()
```

## Top `r params$top_n` Counties % Asian

```{r message = FALSE}
df %>% arrange(desc(asian_pct)) %>% 
  select(state, county, total, asian, asian_pct) %>% 
  head(params$top_n) %>% 
  kable()
```

## Top `r params$top_n` Counties % Native Hawaiian or Other Pacific Islander

```{r message = FALSE}
df %>% arrange(desc(pacific_pct)) %>% 
  select(state, county, total, pacific, pacific_pct) %>% 
  head(params$top_n) %>% 
  kable()
```

## Top `r params$top_n` Counties % Other

```{r message = FALSE}
df %>% arrange(desc(other_pct)) %>% 
  select(state, county, total, other, other_pct) %>% 
  head(params$top_n) %>% 
  kable()
```

## Top `r params$top_n` Counties % Two or More Races

```{r message = FALSE}
df %>% arrange(desc(twoormore_pct)) %>% 
  select(state, county, total, twoormore, twoormore_pct) %>% 
  head(params$top_n) %>% 
  kable()

```